Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-20 paper/ijm.tex
Minor changes to CEE part

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
\fxwarning{SKB: Let's collect first enough material and then decide, but I can start adding here some notes from \cite{Kahanec2009,Kahanec2010}.}
%Start here CEE stuff for now
Comparing the results for Estonia with other Central Eastern European (CEE) countries and former Soviet Union countries is rather difficult due to the lack of data about language skills.
Information about ethnicity is widely available, but no information about individual native and foreign language skills.
For example, the Latvian Labour Force Survey does not (did not) contain questions regarding language skills.
While the language policies in the Baltic states are discussed in \cite{Hogan-Brun2008,Hogan-Brun2007}, the implications for the labour market are not discussed.
\cite{Grasmane2011} use information from a survey conducted at the Latvian University of Agriculture with undergraduate, master, doctoral students to find out  how foreign language skills impact employability.
They find that the better foreign language skills are related to higher willingness to work abroad.
\cite{Alan2015} \fxnote{SKB: Quite detailed article about Russian language proficiency in three former Soviet Union states. Needs further comments from us and find similar articles.}
\cite{Kahanec2009} provide an overview about \enquote{the roles of foreign origin and citizenship in the labor markets of Eastern and Eestern[sic] Member States of the EU} \dots and use the EU Survey of Income and Living Conditions (EU-SILC) dataset.
They\enquote{ mapped and compared the roles of foreign origin and citizenship for labor market performance in Eastern and Western EU Member States.}
Using Oaxaca and Fairlie decompositions, they show that there is native/non-native divide in the national labour markets \dots.
They find that \enquote{both foreign origin and citizenship remain important barriers to non-native labor market integration}
\enquote{In Estonia and Latvia, in particular, employment
	policies seem to be most relevant for non-immigrant non-citizens. Once employed,
	non-nativity may play a less negative role in the labor market for some groups,
	although labor market segmentation remains a significant issue.}
\fxnote{\cite{Kahanec2009} essentially confirms your results in \cite{Leping2008,Toomet2011}}
\cite{Kahanec2010} show that \enquote{the labor market situation of ethnic minorities in the European
	Union. Facing a serious measurement challenge and lacking adequate data, we apply
	several measures of ethnicity and examine various data sources as well as secondary
	evidence. We find significant gaps between ethnic minority and majority populations in
	terms of labor market outcomes.}
%Continue with existing stuff
In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.

>>>>

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
\fxwarning{SKB: Let's collect first enough material and then decide, but I can start adding here some notes from \cite{Kahanec2009,Kahanec2010}.}
%Start here CEE stuff for now
Comparing the results for Estonia with other Central Eastern European (CEE) countries and former Soviet Union countries is rather difficult due to the lack of data about language skills.
Information about ethnicity is widely available, but no information about individual native and foreign language skills.
For example, the Latvian Labour Force Survey does not (did not) contain questions regarding language skills.
While the language policies in the Baltic states are discussed in \cite{Hogan-Brun2008,Hogan-Brun2007}, the implications for the labour market are not discussed.
\cite{Grasmane2011} use information from a survey conducted at the Latvian University of Agriculture with undergraduate, master, doctoral students to find out  how foreign language skills impact employability.
They find that the better foreign language skills are related to higher willingness to work abroad.
\cite{Alan2015} \enquote{During the Soviet era, proficiency in the Russian language was often a ticket to
	attractive employment opportunities in the member republics. Does it still contribute
	to securing employment in the former Soviet republics after two decades of transition?
	Using data from Armenia, Azerbaijan and Georgia in the years 2008–2010,
	this paper demonstrates that Russian language skills remain economically valuable.
	The baseline estimates suggest that Russian language skills increase probability of
	employment by about 6 (males) and 9 (females) percentage points. Our results bear
	important implications for the ongoing debates on language policies in the post-
	Soviet countries.} \fxnote{SKB: Quite detailed article about Russian language proficiency in three former Soviet Union states. Needs further comments from us and find similar articles.}
\cite{Kahanec2009} provide an overview about \enquote{the roles of foreign origin and citizenship in the labor markets of Eastern and Eestern[sic] Member States of the EU} \dots and use the EU Survey of Income and Living Conditions (EU-SILC) dataset.
They\enquote{ mapped and compared the roles of foreign origin and citizenship for labor market performance in Eastern and Western EU Member States.}
Using Oaxaca and Fairlie decompositions, they show that there is native/non-native divide in the national labour markets \dots.
They find that \enquote{both foreign origin and citizenship remain important barriers to non-native labor market integration}
\enquote{In Estonia and Latvia, in particular, employment
	policies seem to be most relevant for non-immigrant non-citizens. Once employed,
	non-nativity may play a less negative role in the labor market for some groups,
	although labor market segmentation remains a significant issue.}
\fxnote{\cite{Kahanec2009} essentially confirms your results in \cite{Leping2008,Toomet2011}}
\cite{Kahanec2010} show that \enquote{the labor market situation of ethnic minorities in the European
	Union. Facing a serious measurement challenge and lacking adequate data, we apply
	several measures of ethnicity and examine various data sources as well as secondary
	evidence. We find significant gaps between ethnic minority and majority populations in
	terms of labor market outcomes.}
%Continue with existing stuff
In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.
